---
title: Dashboard
format: dashboard
---

```{r}
library(tidyverse)
library(sf)
library(ggplot2)
library(dplyr)
library(readxl)

cleaned <- read_csv("/Users/jayson/Documents/stat133/project-5/gerrymandering-jayson-abarca/data/cleaned_precinct_votes.csv")

district_2024 <- read_csv("/Users/jayson/Documents/stat133/project-5/gerrymandering-jayson-abarca/data/g24_sov_by_g24_svprec.csv")

results_2025 <- read_xlsx("/Users/jayson/Documents/stat133/project-5/gerrymandering-jayson-abarca/data/g24-results-by-district.xlsx")

district_2024_sf <- st_read("/Users/jayson/Documents/stat133/project-5/gerrymandering-jayson-abarca/srprec_state_g24_v01_shp/srprec_state_g24_v01_shp.shp")
districts_2025_sf <- st_read("/Users/jayson/Documents/stat133/project-5/gerrymandering-jayson-abarca/AB604.shp")
```

1.  Summary Stats Comparison

```{r}
comparison <- tibble(
  Map = c("2024 District Map", "Proposed 2025 Map"),
  `D Seat Share` = c(sum(district_2024$dem_total > district_2024$rep_total),
                     sum(results_2025$dem_total > results_2025$rep_total)),
  `Mean–Median Score` = c(mm_2024, mm_2025),
  `Efficiency Gap` = c(eg_2024, eg_2025)
)

comparison
```

2.  

```{r}
p1 <- ggplot(districts_2024_sf) +
  geom_sf(aes(fill = factor(cddist))) +
  labs(title = "2024 Congressional Districts")

p2 <- ggplot(districts_2025_sf) +
  geom_sf(aes(fill = factor(new_district))) +
  labs(title = "Proposed AB 604 Districts (2025)")

p1
p2
```

3.  

```{r}
seats <- tibble(
  Map = c("2024", "2025"),
  Dem_Seats = c(sum(district_2024$dem_total > district_2024$rep_total),
                sum(results_2025$dem_total > results_2025$rep_total)),
  Rep_Seats = c(sum(district_2024$rep_total > district_2024$dem_total),
                sum(results_2025$rep_total > results_2025$dem_total))
)

seats
```

4.  

```{r}
bind_rows(
  district_2024 %>% mutate(Plan = "2024"),
  results_2025 %>% mutate(Plan = "2025")
) %>%
  ggplot(aes(x = share_dem, fill = Plan)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Democratic District Vote Share",
       x = "Dem Vote Share", y = "Density")
```

5.  Data Sources:

-   2024 precinct vote data: California Statewide Database

-   2024 district shapefile: Secretary of State

-   AB 604 proposed map shapefile: Assembly Elections Committee

Cleaning:

-   Removed precincts with zero registration or vote totals

-   Converted numeric vote columns using parse_number()

-   Aggregated votes to district level

Reallocation:

-   Used area-weighted interpolation

Metrics:

-   Mean-Median Score: Difference between the average and the median district vote share

-   Efficiency Gap: the differences in wasted votes